---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading libraries to help processing the data
```{r, echo = TRUE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data First check if the file activity.zip was
unzipped, if it was not, then unzip it.
```{r, echo = TRUE}
if (!file.exists('activity.csv')){
    unzip('activity.zip')
}
```
Then read the data, using the colClasses argument to set the Date field as Date.
```{r, echo = TRUE}
data <- read.table('activity.csv',
                    header=TRUE, sep=',',
                    colClasses=c('numeric','Date','numeric'))
```

## What is mean total number of steps taken per day?
First let's calculate the total number of steps taken per day, ignoring NA's.
```{r, echo = TRUE}
total_steps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(total_steps_per_day=sum(steps))
```

Plotting an histogram with the total number of steps.
```{r, echo = TRUE}
hist(total_steps$total_steps_per_day, breaks=50)
```

Calculating the mean of total number of setps taken per day
```{r, echo = TRUE}
mean(total_steps$total_steps_per_day)
```

Calculating the median of total number of setps taken per day
```{r, echo = TRUE}
median(total_steps$total_steps_per_day)
```

## What is the average daily activity pattern?
Plotting  the average number of steps taken for each five minutes
```{r, echo=TRUE}
mean_steps_interval <- data %>%
    select(interval, steps) %>%
    group_by(interval) %>%
    summarize(mean_steps=mean(steps,na.rm = TRUE))
plot(x=mean_steps_interval$interval, y=mean_steps_interval$mean_steps, type='l')
```

The 5-minute interval that has the maximum number of steps, on average across all days is:
```{r, echo=TRUE}
mean_steps_interval %>%
    filter(mean_steps==max(mean_steps)) %>%
    select(interval)
```

## Imputing missing values
The total number of rows with missing is:
```{r, echo=TRUE}
data %>% filter(is.na(steps)) %>% nrow()
```

Filling in all missing values with the mean value for that interval.
```{r, echo=TRUE}
no_na_data <- data
no_na_data[match(mean_steps_interval$interval, data$interval) &
         is.na(data$steps), 'steps'] <- mean_steps_interval$mean_steps
```

Plotting an histogram with the total steps taken each day for the new dataset withou NA
```{r, echo = TRUE}
no_na_total_steps <- no_na_data %>%
    group_by(date) %>%
    summarize(total_steps=sum(steps))
```

Plotting an histogram with the total number of steps.
```{r, echo = TRUE}
hist(no_na_total_steps$total_steps, breaks=50)
```

Calculating the mean of total number of setps taken per day
```{r, echo = TRUE}
mean(no_na_total_steps$total_steps)
```

Calculating the median of total number of setps taken per day
```{r, echo = TRUE}
median(no_na_total_steps$total_steps)
```

Checking if there is any difference on the mean of total number of steps taken per day
```{r, echo=TRUE}
if (mean(no_na_total_steps$total_steps) != mean(total_steps$total_steps_per_day)) {
    mean(no_na_total_steps$total_steps) - mean(total_steps$total_steps_per_day)
    abs(mean(no_na_total_steps$total_steps) - mean(total_steps$total_steps_per_day))/
        mean(total_steps$total_steps_per_day)
}
```

Checking if there is any difference on the median of total number of steps taken per day
```{r, echo=TRUE}
if (median(no_na_total_steps$total_steps) != median(total_steps$total_steps_per_day)) {
    print('The absolute difference on the median is:')
    print(median(no_na_total_steps$total_steps) - median(total_steps$total_steps_per_day))
    print('The percentual change is:')
    print(abs(median(no_na_total_steps$total_steps) - median(total_steps$total_steps_per_day))*100/
        median(total_steps$total_steps_per_day))
}
```

## Are there differences in activity patterns between weekdays and weekends?