---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading libraries to help processing the data
```{r, echo = TRUE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data First check if the file activity.zip was
unzipped, if it was not, then unzip it.
```{r, echo = TRUE}
if (!file.exists('activity.csv')){
    unzip('activity.zip')
}
```
Then read the data, using the colClasses argument to set the Date field as Date.
```{r, echo = TRUE}
data <- read.table('activity.csv',
                    header=TRUE, sep=',',
                    colClasses=c('numeric','Date','numeric'))
```

## What is mean total number of steps taken per day?
First let's calculate the total number of steps taken per day, ignoring NA's.
```{r, echo = TRUE}
total_steps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(total_steps_per_day=sum(steps))
```

Plotting an histogram with the total number of steps.
```{r, echo = TRUE}
hist(total_steps$total_steps_per_day, breaks=50)
```

Calculating the mean of total number of setps taken per day
```{r, echo = TRUE}
mean(total_steps$total_steps_per_day)
```

Calculating the median of total number of setps taken per day
```{r, echo = TRUE}
median(total_steps$total_steps_per_day)
```

## What is the average daily activity pattern?
Plotting  the average number of steps taken for each five minutes


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?